---
title: 'Reproducible Research: Week 2 Assignment'
author: "Avijit"
date: "14^th^ July 2020"
output: 
  html_document: 
    keep_md: yes
---
```{r include=FALSE}
setwd("D:/R working directory")
library(dplyr)
library(data.table)
library(ggplot2)
```

## Loading and processing data
```{r}
if(!file.exists("./activity.csv")){
download.file(url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
              ,destfile = "./activity_data.zip")
unzip(zipfile = "./activity_data.zip")
}
d <- fread("./activity.csv")

```

## Number of steps taken each day

```{r}
bydate <- d %>%
    select(steps,date) %>%
    group_by(date) %>%
    summarise(total_steps = sum(steps)) %>%
    na.omit()

hist(bydate$total_steps,
     xlab = "Total daily Steps",main="Histogram of Total Steps by day",
     breaks = 20)
```

## Mean and median of the total number of steps taken per day
```{r}
mean(bydate$total_steps)
median(bydate$total_steps)
```

## Time series plot of the 5-minute interval and the average number of steps taken, averaged across all days
```{r}
byint <- d %>%
    select(steps, interval) %>%
    na.omit() %>%
    group_by(interval) %>%
    summarise(average_steps = mean(steps))

ggplot(byint, aes(x = interval, y = average_steps)) + geom_line()
```

## 5-minute interval, on average across all the days, that contains the maximum number of steps
```{r}
byint[which.max(byint$average_steps),]
```

## Number of missing values in the dataset
```{r}
length(which(is.na(d$steps)))
```

## Imputing missing values and replacing them with the mean for that 5 -minute interval
```{r}
replacewithmean <- function(x){
    replace(x,is.na(x), mean(x, na.rm = T))
}
byintmean <- d %>%
    group_by(interval) %>%
    mutate(steps = replacewithmean(steps))

bydate_impute <- byintmean %>%
    group_by(date) %>%
    summarise(total_steps = sum(steps))

qplot(total_steps, data = bydate_impute, binwidth = 500,
      xlab='Steps per day', ylab='Frequency',
      main = 'Histogram of the total number of steps taken each day')
```

## Mean and median of imputed and original data
```{r}
summary(bydate_impute)
summary(bydate)
```

## Activity patterns on weekdays and weekends
```{r}
byintmean$date <- as.Date(byintmean$date)
byintmean$day <- weekdays(byintmean$date)
byintmean$weekday <- with(byintmean,
                          ifelse(day == "Saturday"|day == "Sunday","Weekend", "Weekday"))

byweekend <- byintmean %>%
    group_by(weekday,interval) %>%
    summarise(avg = mean(steps))
qplot(interval, avg, data = byweekend, facets = weekday~., color = weekday,
      geom = "line")
```

